% paper/sections/conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

We presented the first systematic study of curriculum domain randomisation
for AUV fluid physics, comparing \CDR{} against Uniform DR, Naive SAC,
and a PID baseline on a held-out test distribution.
Three findings have practical significance for AUV deployment.
First, PID control collapses catastrophically ($100\% \to 3\%$ success)
under modest fluid-parameter shift — a fragility that renders fixed-gain
classical control unsuitable for real ocean deployment without real-time
adaptation.
Second, domain randomisation is essential for reliable transfer:
it eliminates the $\pm [FILL:std_naive]$\% inter-seed variance of
unrobust policies, enabling confident pre-deployment assessment of
policy performance.
Third, \CDR{} produces more energy-efficient policies than Uniform DR
($[FILL:saving]\%$ lower energy per step), a meaningful operational
advantage for battery-constrained AUV missions.

Beyond the specific results, this work establishes energy efficiency
as a meaningful differentiator between DR strategies that achieve
similar success rates.
As AUV RL matures toward real deployment, we argue that mission duration
and thrust economy should join success rate as standard evaluation metrics.
Our open-source Halcyon X4 environment and training pipeline are
released to support this.
Future work includes hardware validation, multi-parameter curriculum
scheduling, and extension to partially observable settings with
degraded sensing.