% paper/sections/discussion.tex

\section{Discussion}
\label{sec:discussion}

\subsection{Why Classical Control Is Insufficient for Robust AUV Deployment}

The $97$-percentage-point collapse of PID performance under out-of-distribution
physics is not a tuning failure — the gains were optimised by exhaustive grid
search on the training distribution.
Rather, it reflects a fundamental structural limitation of fixed-gain linear
control: proportional-integral terms cannot distinguish between position error
caused by inertia (transient, self-correcting) and error caused by persistent
external disturbances such as ocean current (permanent, requiring sustained
compensation).

Under strong tidal currents at $0.60$--$0.80$\,m/s, the integral term
accumulates large corrections, saturates the thruster command, and causes
the AUV to overshoot the goal.
The derivative term damps oscillation near the goal, but with $K_d = 10$
tuned for nominal drag of $0.20$\,kg/m, it is severely underdamped at
test drag values of $0.80$--$1.20$\,kg/m.
The result is a persistent limit cycle that never converges to within the
$0.5$\,m goal threshold.

This analysis generalises beyond our specific implementation: \emph{any}
fixed-gain PID controller designed for nominal AUV conditions will face
analogous degradation in real deployments, where fluid parameters vary
continuously and unpredictably.
Our experiments provide the first precise quantification of this gap for
a $6$-DOF AUV goal-reaching task.

\subsection{Why CDR Produces More Energy-Efficient Policies}

The $[FILL:saving]$\% energy saving of CDR over Uniform DR has a
mechanistic explanation rooted in the curriculum's implicit learning
structure.

During the early phase ($\lambda \approx 0$), the CDR agent trains
exclusively on calm conditions: nominal drag, zero current, nominal buoyancy.
In this regime, the minimum-effort strategy for reaching a goal is
smooth, low-thrust trajectory-following.
The reward's energy penalty ($-0.02\|a_t\|^2$) reinforces this: in calm
conditions, inefficient thrust is penalised without providing any benefit,
so the agent learns to minimise thrust early.

As $\lambda$ increases, the agent encounters mild drag and current
perturbations.
Crucially, it adapts its \emph{existing} low-thrust strategy rather than
developing a new one from scratch.
The curriculum frames each expansion as a perturbation of known behaviour,
not a novel task.

Uniform DR, by contrast, presents all difficulty levels from episode 1.
Under high drag and strong current, aggressive thrust is the only viable
strategy early in training — the agent cannot survive without near-maximum
effort.
This aggressive strategy becomes entrenched in the policy's early weights
and persists even in lower-difficulty conditions where it is unnecessary.

The net result is a systematic difference in \emph{policy texture}: CDR
learns the minimum viable thrust for each condition, while Uniform DR learns
a uniformly aggressive strategy that works everywhere but optimises nowhere.

\subsection{Deployment Implications}

The energy advantage has direct operational significance.
A real AUV mission — pipeline inspection, bathymetric survey, or coral
reef monitoring — is fundamentally battery-constrained.
A $[FILL:saving]$\% reduction in mean thrust consumption translates
approximately linearly to a $[FILL:saving]$\% extension in mission
duration or range, assuming thrust is the dominant power consumer
(which is typical for medium-speed torpedo AUVs~\cite{fossen2011handbook}).

For a 4-hour inspection mission, this represents approximately
$[FILL:duration_min]$\,minutes of additional operational time — sufficient,
in some scenarios, to complete a survey pass that would otherwise require
surfacing and recharging.

We therefore recommend that future AUV RL benchmarks report energy
efficiency alongside success rate as a co-primary metric, particularly
for scenarios where battery endurance constrains mission scope.

\subsection{Generality of CDR}

The CDR algorithm (Algorithm~\ref{alg:cdr}) is not specific to AUV
physics.
Its only requirements are: (1) a parameterisable simulator with identifiable
physics parameters, (2) a binary success signal per episode, and (3) a
reasonable starting range from which the agent can achieve $> \tau^- = 0.40$
success.
These conditions hold broadly across sim-to-real transfer problems.

The simplification relative to ADR~\cite{akkaya2019solving} — expanding
all parameters simultaneously with a single success window rather than
per-parameter expansion buffers — reduces implementation complexity
at the cost of some adaptivity.
For the AUV case, where the six parameters are physically coupled
(current and drag interact multiplicatively), simultaneous expansion
is arguably more appropriate than independent parameter scheduling.
Whether per-parameter expansion yields further improvements is an
open question we leave to future work.

\subsection{Limitations}

\textbf{Simulation-only evaluation.}
This study is limited to simulation.
Hardware validation on a real AUV platform — the critical missing step —
is left to future work.
Tank experiments at $1$--$2$\,m/s current and $0.6$--$1.2$\,kg/m drag
would provide direct evidence for or against the sim-to-real hypothesis.

\textbf{Simplified fluid model.}
Our hydrodynamics follow the standard Fossen~\cite{fossen2011handbook}
formulation but omit turbulence, vortex shedding, wave-induced
disturbances, and thruster interaction effects.
More complex models would increase the training challenge, potentially
accentuating the CDR advantage, but would also require longer training
and more careful tuning.

\textbf{Limited seeds.}
With $n = 3$ seeds, statistical comparisons between closely-matched
conditions (CDR vs Uniform DR) have limited power.
The CDR vs Uniform DR energy difference reaches
$p = [FILL:pval]$ with $d = [FILL:cohend]$;
the effect is practically meaningful but not yet definitively established.
Future work with $n \geq 10$ seeds would provide stronger statistical evidence.

\textbf{Fixed CDR hyperparameters.}
The window size $W = 50$, thresholds $\tau^\pm$, and step sizes $\varepsilon_\pm$
were manually tuned.
Preliminary sensitivity analysis (Appendix~\ref{app:sensitivity}) shows
CDR is robust to moderate variation in $W$, but adaptive meta-learned
thresholds could further improve performance and reduce setup cost.

\textbf{Observation space.}
The current observation does not include sonar, camera, or vision inputs,
limiting applicability to GPS-aided environments.
Extending to partially observable settings with noisy sensing is a direct
and valuable next step.