% paper/sections/introduction.tex

\section{Introduction}
\label{sec:intro}

A PID controller carefully tuned for nominal AUV physics achieves
$100\%$ success rate in simulation — and $3\%$ when the same fluid
parameters shift by a modest, ocean-realistic amount.
This single result, demonstrated quantitatively in
Section~\ref{sec:pid_results}, encapsulates the central challenge of
autonomous underwater vehicle (AUV) deployment: controllers optimised
for nominal conditions are insufficient when drag, buoyancy, and water
current vary with biofouling, salinity gradients, and tidal variation —
as they invariably do in the real ocean.

Reinforcement learning (RL) trained with domain randomisation (DR) offers
a principled solution: by varying physics parameters across training
episodes, policies generalise to the distribution of conditions they
will encounter in deployment~\cite{tobin2017domain,peng2018simtoreal}.
Several recent works have applied SAC and DR to AUV control with
promising results~\cite{arndt2024learning,chu2025marinegym,chaffre2025simtoreal}.
However, standard Uniform DR — sampling uniformly from fixed parameter
ranges throughout training — exposes the agent simultaneously to easy
and hard conditions from the first episode.
We show this produces policies with high inter-seed variance
($60.7\% \pm 27.2\%$ without DR), making deployment reliability uncertain.

A natural improvement is to schedule the difficulty of DR, beginning
with narrow parameter ranges and expanding them as the agent learns to
handle each level.
OpenAI's Automatic Domain Randomisation~\cite{akkaya2019solving} applied
this principle to dexterous manipulation with a robot hand.
However, no prior work has applied curriculum DR to the qualitatively
distinct problem of underwater fluid physics, where parameters are
physically coupled through nonlinear quadratic drag, and no prior work
has evaluated energy efficiency — a metric of direct operational
significance for battery-constrained AUVs.

We present \emph{Curriculum Domain Randomisation} (\CDR{}) for AUV
fluid physics: a performance-gated strategy that starts from narrow
physics ranges and expands them when a rolling success rate window
exceeds a threshold.
\CDR{} is inspired by ADR~\cite{akkaya2019solving} but simplified for
underwater physics — expanding all six parameters simultaneously with a
single shared success window, without per-parameter expansion buffers
or Bayesian models.
We compare \CDR{} against Naive SAC, Uniform DR, and a PID baseline
across $9$ training runs ($3$ conditions $\times$ $3$ seeds $\times$
$10^6$ steps) and evaluate zero-shot transfer to a held-out test
distribution with elevated drag and strong currents.

Our findings are threefold.
First, PID control collapses catastrophically under fluid-parameter
distribution shift ($100\% \to 3\%$ success), establishing a
practical baseline for the value of physics-aware training.
Second, any form of DR eliminates the $\pm 27.2\%$ deployment variance
of unrobust policies, confirming that distribution coverage is
necessary for reliable transfer.
Third, \CDR{} produces policies that consume $[FILL:saving]\%$ less
thrust energy per step than Uniform DR at comparable success rates —
a practically meaningful saving for battery-constrained real AUVs
where mission duration determines operational scope.

\textbf{Contributions:}
\begin{enumerate}
\item The first systematic comparison of curriculum vs uniform DR
  strategies for AUV fluid physics, demonstrating that CDR produces
  more energy-efficient policies at comparable transfer success rates.
\item A quantitative benchmark showing that PID control collapses from
  $100\%$ to $3\%$ success under out-of-distribution fluid parameters —
  establishing the practical necessity of physics-aware training for
  real AUV deployment.
\item An open-source MuJoCo AUV environment (\textit{Halcyon X4}) with
  nine randomisable parameters (six physics $+$ three sensor noise),
  released for reproducible underwater robotics research at
  \url{https://github.com/Itzlimon22/rl_robotics}.
\end{enumerate}