% paper/sections/related_work.tex

\section{Related Work}
\label{sec:related}

\subsection{Domain Randomisation for Sim-to-Real Transfer}

Domain randomisation (DR) trains policies across a distribution of simulated
environments, with the hypothesis that the real world falls within the
training distribution~\cite{tobin2017domain}.
Tobin et al.~\cite{tobin2017domain} demonstrated visual DR for robotic
object localisation, achieving zero-shot sim-to-real transfer purely from
randomised rendered images.
Peng et al.~\cite{peng2018simtoreal} extended DR to dynamics randomisation
for legged locomotion, randomising mass, friction, and joint parameters.

Adaptive approaches to DR — adjusting randomisation ranges during training —
have shown further gains.
OpenAI's Automatic Domain Randomisation (ADR)~\cite{akkaya2019solving}
expands parameter boundaries when a per-parameter performance buffer
indicates competence, enabling the learning of dexterous manipulation with
132 randomised parameters.
DORAEMON~\cite{tiboni2024doraemon} frames adaptive DR as entropy maximisation
subject to a success rate constraint, using Bayesian optimisation to select
parameter distributions.

Our CDR algorithm adapts the ADR concept to underwater fluid physics,
expanding all six parameters simultaneously using a single shared success
window rather than per-parameter buffers.
This simplification reduces the hyperparameter count and is appropriate
for the AUV setting where parameters are physically coupled.
Unlike ADR and DORAEMON, which target manipulation on ground-based systems,
we apply curriculum DR to the qualitatively distinct parameter space of
hydrodynamic forces, and we evaluate energy efficiency as a primary metric.

\subsection{Reinforcement Learning for AUV Control}

Deep RL has been applied to AUV station-keeping~\cite{carlucho2018},
depth and heading control~\cite{shi2020survey}, and 3D navigation.
Arndt et al.~\cite{arndt2024learning} demonstrated zero-shot sim-to-real
transfer of a $6$-DOF AUV controller using SAC with Uniform DR on two
physical parameters (centre of buoyancy and displaced volume), achieving
performance comparable to hand-tuned PID on a real vehicle.
Chaffre et al.~\cite{chaffre2025simtoreal} combined SAC with
bio-inspired experience replay and enhanced Uniform DR for AUV
station-keeping under current disturbances, with physical tank validation.

MarineGym~\cite{chu2025marinegym} provides the most comprehensive
underwater RL benchmark to date: GPU-accelerated simulation via
NVIDIA Isaac Sim with five AUV models, three task types, and a
Uniform DR toolkit.
However, MarineGym does not study curriculum scheduling of DR,
does not evaluate energy efficiency, and uses proprietary NVIDIA
simulation rather than the open-source MuJoCo we employ.

Neither Arndt et al.~\cite{arndt2024learning} nor
Chaffre et al.~\cite{chaffre2025simtoreal} study curriculum DR or
measure energy efficiency.
No prior work quantifies the PID fragility gap under fluid-parameter
distribution shift, which we show to be a $97$-percentage-point collapse.

\subsection{Curriculum Learning and Self-Paced DR}

Curriculum learning~\cite{bengio2009curriculum} — presenting training
examples in order from easy to hard — has been shown to improve
generalisation and convergence in supervised and reinforcement learning.
Self-paced RL methods~\cite{klink2020selfpaced} adapt task difficulty
to the current learner's capability, using contextual MDPs or target
distributions.

In the DR context, curriculum approaches that expand randomisation
ranges over time have been explored primarily for manipulation and
locomotion.
Our work is, to our knowledge, the first to apply performance-gated
curriculum DR to underwater fluid physics, and the first to demonstrate
that curriculum scheduling produces policies with systematically lower
energy consumption than Uniform DR — a result with direct practical
significance for battery-constrained AUV deployment.