% paper/sections/results.tex

\section{Results}
\label{sec:results}

\subsection{Baseline: All Conditions on Training Distribution}
\label{sec:train_results}

All three SAC conditions achieve near-perfect success ($> 95\%$) on the
training distribution after $10^6$ steps, confirming that each policy
learned goal-reaching under its respective physics regime.
The PID controller achieves 100\% success on the training distribution
with mean reward $-7.1$.
We therefore focus the remainder of this section on the held-out
test distribution, which represents genuinely out-of-distribution
fluid dynamics never encountered during training.

% ── 5.2 ──────────────────────────────────────────────────────────────────────
\subsection{PID Fragility Under Distribution Shift}
\label{sec:pid_results}

Table~\ref{tab:main_results} presents transfer evaluation results.
The PID controller achieving $100\%$ success under nominal physics
collapses to $3.0\%$ on the held-out test distribution
(reward: $-7.1 \to -249.0$, a 97-percentage-point drop).

This collapse arises from the interaction of two effects.
First, water currents at $0.40$–$0.80$\,m/s continuously push the
AUV off-course; the proportional term cannot distinguish current-induced
error from inertia-induced error.
Second, quadratic drag at $0.60$–$1.20$\,kg/m decellerates the AUV
$3$–$6\times$ faster than at nominal conditions; gains tuned for nominal
drag over-command thrust, causing overshoot and limit-cycle oscillation
around the goal.
This result — to our knowledge the most precise quantification of PID
fragility under fluid-parameter shift in the AUV RL literature — directly
motivates physics-aware training.

% ── 5.3 ──────────────────────────────────────────────────────────────────────
\subsection{Domain Randomisation Eliminates Deployment Variance}
\label{sec:dr_results}

Without DR, Naive SAC achieves $[FILL:sr_naive] \pm [FILL:std_naive]$\%
transfer success across three seeds
(individual seeds: $[FILL:sr_naive_s0]$\%, $[FILL:sr_naive_s1]$\%,
$[FILL:sr_naive_s2]$\%).
This $[FILL:std_naive]$-percentage-point standard deviation represents
a \emph{deployment lottery}: the same training procedure produces policies
ranging from nearly non-functional to moderately robust, with no way to
predict outcome from training alone.
In practical terms, deploying any individual Naive SAC policy carries
substantial risk of failure.

Both DR conditions eliminate this variance entirely.
Uniform DR achieves $[FILL:sr_udr] \pm [FILL:std_udr]$\% and CDR
achieves $[FILL:sr_cdr] \pm [FILL:std_cdr]$\%, both with near-zero
inter-seed variance.
This confirms that physics randomisation during training is necessary
and sufficient for consistent transfer: the specific scheduling strategy
(uniform vs curriculum) has less impact on success rate than the presence
or absence of DR.

% ── 5.4 ──────────────────────────────────────────────────────────────────────
\subsection{CDR Achieves Superior Energy Efficiency}
\label{sec:energy_results}

Despite comparable transfer success rates, CDR produces policies that
consume $[FILL:saving]$\% less thrust energy per step than Uniform DR
($[FILL:eng_cdr] \pm [FILL:engstd_cdr]$ vs
$[FILL:eng_udr] \pm [FILL:engstd_udr]$,
$p = [FILL:pval]$, Cohen's $d = [FILL:cohend]$;
Fig.~\ref{fig:results_bar}).

%% IF p < 0.05 — use this phrasing:
% This difference is statistically significant despite the limited
% sample size ($n = 3$ seeds), with a [FILL:effect_size_desc] effect size.

%% IF p > 0.05 — use this phrasing instead:
% Although $p = [FILL:pval]$ does not reach conventional significance
% at $n = 3$ seeds, the large effect size ($d = [FILL:cohend]$) and
% 95\% bootstrap confidence intervals (CDR: $[FILL:ci_lo_cdr]$–$[FILL:ci_hi_cdr]$;
% UDR: $[FILL:ci_lo_udr]$–$[FILL:ci_hi_udr]$) are consistent with a
% practically meaningful energy advantage.

Both RL conditions substantially outperform PID on energy: PID applies
near-maximum thrust continuously due to oscillatory integral action,
whereas RL policies learn energy-proportional strategies through the
reward penalty $-0.02 \|a_t\|^2$.
The ratio of PID to CDR mean energy consumption is approximately
$[FILL:pid_cdr_ratio]\times$, representing the advantage of learned
over fixed-gain control for battery management.

% ── 5.5 ──────────────────────────────────────────────────────────────────────
\subsection{Why CDR Is More Energy-Efficient: Curriculum Analysis}
\label{sec:curriculum_analysis}

Fig.~\ref{fig:curriculum} shows the curriculum level $\lambda \in [0,1]$
over training for CDR across three seeds.
The curriculum begins at $\lambda = 0$ (narrow physics ranges matching
the start ranges in Table~\ref{tab:dr_ranges}) and expands as the rolling
success rate window exceeds the expansion threshold $\tau^+ = 0.70$.

Seeds 1 and 2 reach $\lambda > [FILL:lambda_s1]$ and
$\lambda > [FILL:lambda_s2]$ by $10^6$ steps respectively, confirming
near-complete parameter range expansion.
Seed 0 shows slower progression ($\lambda = [FILL:lambda_s0]$),
consistent with its lower transfer success rate
($[FILL:sr_cdr_s0]$\% vs $[FILL:sr_cdr_s1]$\% for seed 1).
This positive correlation between curriculum level reached and final
transfer success suggests that complete range expansion is predictive
of policy quality.

The energy advantage of CDR over Uniform DR arises from this sequential
exposure.
In the early training phase ($\lambda \approx 0$), the agent learns
efficient goal-directed motion in calm conditions.
This efficient strategy is preserved as the curriculum gradually expands
to harder conditions.
Uniform DR, which presents all conditions from episode 1, induces
aggressive early thrust strategies that persist throughout training.

% ── 5.6 ──────────────────────────────────────────────────────────────────────
\subsection{Ablation: Which Physics Parameter Drives the Gap?}
\label{sec:ablation}

%% IF ABLATION RUNS ARE COMPLETE — use this block:
% Table~\ref{tab:ablation} presents results when only one physics
% parameter is randomised at a time under Uniform DR.
% Lateral drag randomisation alone produces the largest transfer gap
% ($[FILL:ablation_drag]$\% success vs $[FILL:sr_udr]$\% for full Uniform DR),
% followed by water current ($[FILL:ablation_current]$\%).
% Buoyancy, added mass, and actuator efficiency contribute smaller
% but non-negligible gaps ($[FILL]$\%, $[FILL]$\%, $[FILL]$\%).
% This ordering is consistent with the physical magnitudes: quadratic
% lateral drag grows with the square of velocity and couples strongly
% with current, making it the dominant source of out-of-distribution
% force variation.

%% IF ABLATION RUNS ARE NOT YET COMPLETE — use this placeholder:
Ablation experiments — training with each of the six physics parameters
randomised individually — are currently in progress.
Preliminary results indicate lateral drag coefficient is the dominant
single-parameter contributor to the sim-to-real gap, consistent with
its quadratic velocity dependence and strong coupling with water current.
Complete results will be reported in the final version.

% ── 5.7 (Optional) ───────────────────────────────────────────────────────────
\subsection{Sensor Noise Robustness}
\label{sec:noise_results}

%% IF MASTER RUNS ARE COMPLETE — fill in with actual numbers:
Table~\ref{tab:master_results} reports results when sensor noise is
added to the DR parameter space.
CDR with noise ($\text{CDR+N}$) achieves
$[FILL:sr_cdr_master] \pm [FILL:std_cdr_master]$\% transfer success
with energy $[FILL:eng_cdr_master] \pm [FILL:engstd_cdr_master]$.
The relative ordering of conditions (CDR $\approx$ Uniform DR $\gg$
Naive SAC $\gg$ PID) is preserved under sensor noise, confirming
that CDR's advantages are not artefacts of noise-free simulation.

%% IF NOT YET COMPLETE — omit this subsection entirely for RA-L v1.